Structural Divergence Analysis

### 2016-11-04 analysis-structure-window.Rmd
cambios 04/11/16

dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.sum.dif.dist.csv", sep = ""))

dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.sum.dif.dist.csv", sep = "")) #cambiar

dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.sum.dif.dist.csv", sep = ""))
```

@@

siteComparison.all.datasets = ddply(siteComparison.all.datasets, c("dataset", "protein"), mutate,
          "n.SD" = dri2/mean(dri2, na.rm = T),
          "z.SD" = vscale(dri2),
          "RSD" = (dri2),
          "n.RSD" = RSD/mean(RSD, na.rm = T),
          "z.RSD" = vscale(RSD))
```

>>>>

dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.windows.rot.csv", sep = ""))

dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.rot.csv", sep = "")) #cambiar

dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.rot.csv", sep = ""))
```

@@

siteComparison.all.datasets = ddply(siteComparison.all.datasets, c("dataset", "protein"), mutate,
          "n.SD" = dri2/mean(dri2, na.rm = T),
          "z.SD" = vscale(dri2),
          "RSD" = sqrt(dri2), 
          "n.RSD" = RSD/mean(RSD, na.rm = T),
          "z.RSD" = vscale(RSD))
```

### 2016-11-04 analysis-structure-window2.Rmd
cambios 04/11/16

@@

cor.df = data.frame("comparison" = c("(exp - mut) vs min.da.CA", "(exp - mut) vs min.da.CM"),
                    "R" = c(r.min.da.CA, r.min.da.CM)) 
knitr::kable(cor.df, digits = 2)

## find active sites
d$label = as.character(d$site)
active.sites = d$site[d$shell == 0]

## start with a print of smoothed profiles
plot(smooth(z.RSD.exp), col = "blue", type = "l")
lines(smooth(z.RSD.strong), col = "red")
lines(smooth(z.RSD.mut), col = "green")
abline(v = active.sites, col = "grey")

## plot mean profiles with deviations
p = ggplot(d,aes(x = site, y = mean.z.RSD, col = dataset, fill = dataset)) +
    geom_line() +
    geom_ribbon(aes(ymin = mean.z.RSD - f95 * se.z.RSD, ymax = mean.z.RSD + se.z.RSD), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, col = "grey", show.legend = F) +
    labs(x = "site", y = "structural divergence(mean z.RSD)") +
    coord_cartesian(ylim = c(-1.5, 2.8)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

## p + theme(legend.position = "bottom", legend.title = element_blank())
p + theme(legend.position = "bottom")

```{r mean-n-RSD-vs-site}
# comparison of n.RSD profiles with functional sites

## dat with means and standard errors
d = ddply(dat,c("dataset","site","AA","shell"), function(x) {
                mean.n.RSD = mean(x$n.RSD, na.rm = T)
                ndata = sum(!is.na(x$n.RSD))
                se.n.RSD = sd(x$n.RSD, na.rm = T) / sqrt(ndata)
                data.frame(mean.n.RSD, ndata, se.n.RSD)
})

## comparison of means with the minimum distance to the functional sites
n.RSD.exp = d$mean.n.RSD[d$dataset == "exp"]
n.RSD.strong = d$mean.n.RSD[d$dataset == "mut+sel"]
n.RSD.mut = d$mean.n.RSD[d$dataset == "mut"]

plot(y = n.RSD.exp - n.RSD.mut, x = m.min.da.CM.ca$min.da.ca, xlab = "minimal distance to active site (between CAs)")
plot(y = n.RSD.exp - n.RSD.mut, x = m.min.da.CM.ca$min.da.CM, xlab = "minimal distance to active site (between CMs)")

r.min.da.CA = cor((n.RSD.exp - n.RSD.mut), m.min.da.CM.ca$min.da.ca)
r.min.da.CM = cor((n.RSD.exp - n.RSD.mut), m.min.da.CM.ca$min.da.CM)

cor.df = data.frame("comparison" = c("(exp - mut) vs min.da.CA", "(exp - mut) vs min.da.CM"),
                    "R" = c(r.min.da.CA, r.min.da.CM)) 
knitr::kable(cor.df, digits = 2)

## start with a print of smoothed profiles
plot(smooth(n.RSD.exp), col = "blue", type = "l")
lines(smooth(n.RSD.strong), col = "red")
lines(smooth(n.RSD.mut), col = "green")
abline(v = active.sites)

## plot mean profiles with deviations
p = ggplot(d,aes(x = site, y = mean.n.RSD, col = dataset, fill = dataset)) +
    geom_line() +
    geom_ribbon(aes(ymin = mean.n.RSD - f95 * se.n.RSD, ymax = mean.n.RSD + se.n.RSD), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, col = "grey", show.legend = F) +
    labs(x = "site", y = "structural divergence(mean n.RSD)") +
    coord_cartesian(ylim = c(-1.5, 2.8)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

## p + theme(legend.position = "bottom", legend.title = element_blank())
p + theme(legend.position = "bottom")

@@

ggplot(dat, aes(x = site, y = mean.z.RSD, col = dataset)) + geom_line() +
    geom_text(data = subset(dat, dataset == "exp" & mean.z.RSD < -.5), aes(x = site, y = mean.z.RSD, label = site), nudge_y = -.1, size = 1.8)+
    geom_vline(xintercept = active.sites, col = "grey", show.legend = F)

## calculate correlations
r.exp.mut = cor(dat.exp$mean.z.RSD, dat.mut$mean.z.RSD)
r.exp.strong = cor(dat.exp$mean.z.RSD, dat.strong$mean.z.RSD)
r.mut.strong = cor(dat.mut$mean.z.RSD, dat.strong$mean.z.RSD)
cor.df = data.frame("comparison" = c("exp vs mut", "exp vs strong", "mut vs strong"),
                    "R" = c(r.exp.mut, r.exp.strong, r.mut.strong))
knitr::kable(cor.df, digits = 2)

# norm
dat = siteComparison.all.datasets

## dat with means and standard errors
dat = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.n.RSD = mean(x$n.RSD, na.rm = T)
                ndata = sum(!is.na(x$n.RSD))
                se.n.RSD = sd(x$n.RSD, na.rm = T) / sqrt(ndata)
                data.frame(mean.n.RSD, ndata, se.n.RSD)
})

ggplot(dat, aes(x = site, y = mean.n.RSD, col = dataset)) + geom_line() +
    geom_text(data = subset(dat, dataset == "exp" & mean.n.RSD < -.5), aes(x = site, y = mean.n.RSD, label = site), nudge_y = -.1, size = 1.8)+
    geom_vline(xintercept = active.sites, col = "grey", show.legend = F)

## calculate correlations
r.exp.mut = cor(dat.exp$mean.n.RSD, dat.mut$mean.n.RSD)
r.exp.strong = cor(dat.exp$mean.n.RSD, dat.strong$mean.n.RSD)
r.mut.strong = cor(dat.mut$mean.n.RSD, dat.strong$mean.n.RSD)
cor.df = data.frame("comparison" = c("exp vs mut", "exp vs strong", "mut vs strong"),
                    "R" = c(r.exp.mut, r.exp.strong, r.mut.strong))
knitr::kable(cor.df, digits = 2)
```

>>>>

## find active sites
d$label = as.character(d$site)
active.sites = d$site[d$shell == 0]
contacts.active.sites = d$site[d$shell == 1]

## prepare plots with mean profiles and deviations
p = ggplot(d,aes(x = site, y = mean.z.RSD, col = dataset, fill = dataset)) +
    geom_line() +
    geom_ribbon(aes(ymin = mean.z.RSD - f95 * se.z.RSD, ymax = mean.z.RSD + se.z.RSD), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, show.legend = F) +
    labs(x = "site", y = "structural divergence(mean z.RSD)") +
    coord_cartesian(ylim = c(-1.5, 2.8)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

## plot superimposed plots
p + theme(legend.position = "bottom")

@@

cor.df = data.frame("comparison" = c("(exp - mut) vs min.da.CA", "(exp - mut) vs min.da.CM"),
                    "R" = c(r.min.da.CA, r.min.da.CM)) 
knitr::kable(cor.df, digits = 2)
```

@@

ggplot(dat, aes(x = site, y = mean.z.RSD, col = dataset)) + geom_line() +
    geom_text(data = subset(dat, dataset == "exp" & mean.z.RSD < -.5), aes(x = site, y = mean.z.RSD, label = site), nudge_y = -.1, size = 1.8)+
    geom_vline(xintercept = active.sites, col = "grey", show.legend = F) +
    geom_vline(xintercept = contacts.active.sites, col = "black", show.legend = F)

## calculate correlations
r.exp.mut = cor(dat.exp$mean.z.RSD, dat.mut$mean.z.RSD)
r.exp.strong = cor(dat.exp$mean.z.RSD, dat.strong$mean.z.RSD)
r.mut.strong = cor(dat.mut$mean.z.RSD, dat.strong$mean.z.RSD)
cor.df = data.frame("comparison" = c("exp vs mut", "exp vs strong", "mut vs strong"),
                    "R" = c(r.exp.mut, r.exp.strong, r.mut.strong))
knitr::kable(cor.df, digits = 2)
```

### 2016-11-04 analysis-structure-window3.Rmd
cambios 04/11/16

## profiles to compare
local.score.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.local.score.csv", sep = ""))
local.score.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.local.score.csv", sep = ""))
local.score.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.local.score.csv", sep = ""))
```

@@

```{r read_data}
# read data
## exp
local.score.exp = read.csv(local.score.exp.fname, header = T)
site = site.info$site
local.score.exp = local.score.exp[1:length(protein.exp), site]
rownames(local.score.exp) = protein.exp

## mut
local.score.mut = read.csv(local.score.mut.fname, header = T)
site = site.info$site
local.score.mut = local.score.mut[1:length(protein.exp), site]
rownames(local.score.mut) = paste("mut", rownames(local.score.mut), sep = "")

## strong
local.score.strong = read.csv(local.score.strong.fname, header = T)
site = site.info$site
local.score.strong = local.score.strong[1:length(protein.exp), site]
rownames(local.score.strong) = paste("strong", rownames(local.score.strong), sep = "")
```

@@

```{r prepare_site_data}
# prepare the data
## exp
tlocal.score.exp = as.data.frame(t(as.matrix(local.score.exp)))
tlocal.score.exp = cbind(site.info, tlocal.score.exp)
siteComparison.exp = melt(tlocal.score.exp, id.vars = names(site.info), variable.name = "protein", value.name = "local.score")

## mut
tlocal.score.mut = as.data.frame(t(as.matrix(local.score.mut)))
tlocal.score.mut = cbind(site.info, tlocal.score.mut)
siteComparison.mut = melt(tlocal.score.mut, id.vars = names(site.info), variable.name = "protein", value.name = "local.score")

## strong
tlocal.score.strong = as.data.frame(t(as.matrix(local.score.strong)))
tlocal.score.strong = cbind(site.info, tlocal.score.strong)
siteComparison.strong = melt(tlocal.score.strong, id.vars = names(site.info), variable.name = "protein", value.name = "local.score")

@@

siteComparison.all.datasets = ddply(siteComparison.all.datasets, c("dataset", "protein"), mutate,
          "local.score" = (local.score), 
          "n.local.score" = local.score/mean(local.score, na.rm = T),
          "z.local.score" = vscale(local.score))
```

### Mean z.local.score vs site

To measure structural divergence we use z.local.score profiles, which are the best for most families.

The following plot compares the mean z.local.score profiles, including standard errors (with .999 confidence level for the mean)
*Warning: the effective number of data should be smaller than the actual data, especially for the experimental dataset because of several factors. We need to deal with this for quantitative analysis. Probably we'll need to use bootstrapping to calculate standard errors non-parametrically*

```{r mean_se_z.local.score_vs_site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.local.score profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("strong" = "mut+sel"))

## dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.z.local.score = mean(x$z.local.score, na.rm = T)
                ndata = sum(!is.na(x$z.local.score))
                se.z.local.score = sd(x$z.local.score, na.rm = T) / sqrt(ndata)
                data.frame(mean.z.local.score, ndata, se.z.local.score)
})

@@

## prepare plots with mean profiles and deviations
p = ggplot(d, aes(x = site, y = mean.z.local.score, col = dataset, fill = dataset)) +
    geom_line() +
    geom_ribbon(aes(ymin = mean.z.local.score - f95 * se.z.local.score, ymax = mean.z.local.score + se.z.local.score), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, show.legend = F) +
    geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) +
    labs(x = "site", y = "structural divergence(mean z.local.score)") +
    coord_cartesian(ylim = c(-1.5, 2.8)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

@@

```{r_comparison_mean_z.local.score_vs_site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.local.score profiles
ggplot(d, aes(x = site, y = mean.z.local.score, col = dataset)) +
  geom_line() +
  geom_text(data = subset(d, dataset == "exp" & mean.z.local.score < -.5), aes(x = site, y = mean.z.local.score, label = site), nudge_y = -.1, size = 1.8)+
  geom_vline(xintercept = active.sites, show.legend = F) +
  geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) + 
  scale_colour_manual(name = NULL, values = myColors)

@@

r.exp.mut = cor(d.exp$mean.z.local.score, d.mut$mean.z.local.score)
r.exp.strong = cor(d.exp$mean.z.local.score, d.strong$mean.z.local.score)
r.mut.strong = cor(d.mut$mean.z.local.score, d.strong$mean.z.local.score)

@@

# comparison of means.z.scores with the minimum distance to the functional sites
plot(y = d.exp$mean.z.local.score - d.mut$mean.z.local.score, x = m.min.da.CM.CA$min.da.ca, xlab = "minimal distance to active site (between CAs)", ylab = "exp.mean.z.local.score - mut.mean.z.local.score")
plot(y = d.exp$mean.z.local.score - d.mut$mean.z.local.score, x = m.min.da.CM.CA$min.da.CM, xlab = "minimal distance to active site (between CMs)", ylab = "exp.mean.z.local.score - mut.mean.z.local.score")

## calculate correlations between profiles
r.min.da.CA = cor((d.exp$mean.z.local.score - d.mut$mean.z.local.score), m.min.da.CM.CA$min.da.ca)
r.min.da.CM = cor((d.exp$mean.z.local.score - d.mut$mean.z.local.score), m.min.da.CM.CA$min.da.CM)

>>>>

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.windows.rot.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.rot.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.rot.csv", sep = ""))
```

@@

```{r read_data}
# read data
## exp
dri2.exp = read.csv(dri2.exp.fname, header = T)
site = site.info$site
dri2.exp = dri2.exp[1:length(protein.exp), site]
rownames(dri2.exp) = protein.exp

## mut
dri2.mut = read.csv(dri2.mut.fname, header = T)
site = site.info$site
dri2.mut = dri2.mut[1:length(protein.exp), site]
rownames(dri2.mut) = paste("mut", rownames(dri2.mut), sep = "")

## strong
dri2.strong = read.csv(dri2.strong.fname, header = T)
site = site.info$site
dri2.strong = dri2.strong[1:length(protein.exp), site]
rownames(dri2.strong) = paste("strong", rownames(dri2.strong), sep = "")
```

@@

```{r prepare_site_data}
# prepare the data
## exp
tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## mut
tdri2.mut = as.data.frame(t(as.matrix(dri2.mut)))
tdri2.mut = cbind(site.info, tdri2.mut)
siteComparison.mut = melt(tdri2.mut, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

## strong
tdri2.strong = as.data.frame(t(as.matrix(dri2.strong)))
tdri2.strong = cbind(site.info, tdri2.strong)
siteComparison.strong = melt(tdri2.strong, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

@@

siteComparison.all.datasets = ddply(siteComparison.all.datasets, c("dataset", "protein"), mutate,
          "n.SD" = dri2/mean(dri2, na.rm = T),
          "z.SD" = vscale(dri2),
          "RSD" = sqrt(dri2), 
          "n.RSD" = RSD/mean(RSD, na.rm = T),
          "z.RSD" = vscale(RSD))
```

### Mean z.RSD vs site

To measure structural divergence we use z.RSD profiles, which are the best for most families.

The following plot compares the mean z.RSD profiles, including standard errors (with .999 confidence level for the mean)
*Warning: the effective number of data should be smaller than the actual data, especially for the experimental dataset because of several factors. We need to deal with this for quantitative analysis. Probably we'll need to use bootstrapping to calculate standard errors non-parametrically*

```{r mean_se_z.RSD_vs_site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles with functional sites
## prepare dat
dat = siteComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("strong" = "mut+sel"))

## dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                ndata = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(ndata)
                data.frame(mean.z.RSD, ndata, se.z.RSD)
})

@@

## prepare plots with mean profiles and deviations
p = ggplot(d, aes(x = site, y = mean.z.RSD, col = dataset, fill = dataset)) +
    geom_line() +
    geom_ribbon(aes(ymin = mean.z.RSD - f95 * se.z.RSD, ymax = mean.z.RSD + se.z.RSD), alpha = .4, show.legend = F) +
    geom_vline(xintercept = active.sites, show.legend = F) +
    geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) +
    labs(x = "site", y = "structural divergence(mean z.RSD)") +
    coord_cartesian(ylim = c(-1.5, 2.8)) +
    scale_colour_manual(name = NULL, values = myColors) + 
    scale_fill_manual(name = NULL, values = myColors)

@@

```{r_comparison_mean_z.RSD_vs_site, fig.cap = "Black lines are the active sites and grey lines are their contacts"}
# comparison of z.RSD profiles
ggplot(d, aes(x = site, y = mean.z.RSD, col = dataset)) +
  geom_line() +
  geom_text(data = subset(d, dataset == "exp" & mean.z.RSD < -.5), aes(x = site, y = mean.z.RSD, label = site), nudge_y = -.1, size = 1.8)+
  geom_vline(xintercept = active.sites, show.legend = F) +
  geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) + 
  scale_colour_manual(name = NULL, values = myColors)

@@

r.exp.mut = cor(d.exp$mean.z.RSD, d.mut$mean.z.RSD)
r.exp.strong = cor(d.exp$mean.z.RSD, d.strong$mean.z.RSD)
r.mut.strong = cor(d.mut$mean.z.RSD, d.strong$mean.z.RSD)

@@

# comparison of means.z.scores with the minimum distance to the functional sites
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.ca, xlab = "minimal distance to active site (between CAs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")
plot(y = d.exp$mean.z.RSD - d.mut$mean.z.RSD, x = m.min.da.CM.CA$min.da.CM, xlab = "minimal distance to active site (between CMs)", ylab = "exp.mean.z.RSD - mut.mean.z.RSD")

## calculate correlations between profiles
r.min.da.CA = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.ca)
r.min.da.CM = cor((d.exp$mean.z.RSD - d.mut$mean.z.RSD), m.min.da.CM.CA$min.da.CM)

### 2016-11-04 analysis-structure3.Rmd
cambios 04/11/16

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.windows.rot.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.rot.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.windows.rot.csv", sep = ""))
```

>>>>

## profiles to compare
dri2.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.norm.dr.squarei.csv", sep = ""))
dri2.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
dri2.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.norm.dr.squarei.csv", sep = ""))
```
